\section{Conclusions}
\label{sec:conclusions}

This work provides a connected simulation and digitization chain for the UBT,
from a detailed standalone optical model to a computationally inexpensive
FairShip response.  The standalone Geant4 application supports two
configurable scintillator geometries: the \(40\times40\times10~\mathrm{mm^3}\)
tile representing the \(4\times4~\mathrm{cm^2}\) constituent type and the
\(20\times20\times5~\mathrm{mm^3}\) tile representing the
\(2\times2~\mathrm{cm^2}\) type.  Both use the same
\(6\times6~\mathrm{mm^2}\) photosensor, allowing the tile-size dependence of
light collection and ADC response to be compared without changing the sensor
acceptance.

The standalone output records event-level energy deposition, scintillation
yield, sensor photon counts, photoelectron counts, charge, ADC response, hit
position, and selected photoelectron arrival times.  The accompanying
analysis chain produces charge and light-yield distributions, response
correlations, spatial color maps, and CSV tables of mean ADC/MeV in
\(1\times1~\mathrm{mm^2}\) position bins.  These tables preserve the most
important average position dependence of the optical calculation in a form
that can be evaluated quickly during full FairShip production.  Separate maps
are retained for the 20 mm and 40 mm tiles.

In FairShip, a common YAML map defines 225 physical
\(40\times40~\mathrm{cm^2}\) transport regions and specifies whether each
region represents 2 cm or 4 cm constituent segmentation.  The final map has
102 small-tile regions and 123 large-tile regions.  Its full-height green
sides each contain three parent columns, while each side of the three-row
central neck extends through two additional green columns; the innermost
neck is a five-column blue core.  The constituent tiles themselves are not
individual geometry volumes.  For every
\texttt{UpstreamTaggerPoint}, digitization calculates the appropriate
implicit tile center and returns that center as the reconstructed transverse
position.  The original exact Monte Carlo position is retained long enough
to calculate the local coordinates within the inferred tile.  FairShip then
multiplies the deposited energy by the nearest populated ADC/MeV map value,
stores the rounded ADC count, and marks the hit as triggered when the result
is at least 100 counts.  This procedure captures the mean optical response
without generating or transporting optical photons inside FairShip.

The charge-response implementation and its calibration workflow are usable
in their present simplified form, but several effects remain outside the
model.  The sensor efficiency is wavelength independent, the electronics
response omits gain fluctuations, noise, shaping, and saturation, and the
FairShip map lookup is deterministic and uses nearest-neighbor position bins.
Multiple points in one inferred tile are also digitized separately rather
than merged before threshold evaluation.  These limitations should be
considered when interpreting absolute ADC spectra or trigger efficiencies.

The timing analysis requires particular caution.  Absolute photoelectron
arrival times and first-photon-referenced threshold times are available as
diagnostic studies, including Gaussian-fit timing resolutions.  However, the
current simulated distributions do not agree with the real tile data and are
not yet validated for detector digitization.  The present FairShip timing
model therefore relies on the resolution measured in real data and applies it
as a smearing term; it does not import the standalone simulated timing
distribution.  Further comparison with data, tuning of the optical and sensor
models, and validation of the two-tile timing observable are required before
the simulated timing response can replace that data-driven treatment.

